### 2016-10-11 ms/ms/ms.tex
various bug fixes

First, to improve efficiency, filter over latent states
to calculate model likelihoods in order to limit MCMC sampling to
top-level parameters. We then use two computational approaches to
improve the efficiency of this model 1) dynamic blocking of the
parameters \citep{turek2016efficient}, and 2) a custom MCMC
specification where a slice sampler is used for all parameters.

@@

To improve the efficiency of this model, we first filtered over latent
states to calculate model likelihoods in order to limit MCMC sampling
to top-level parameters. We also vectorized all calculations that
would have require for loops in BUGS or JAGS. We then tried two
approaches to speed sampling of the top-level parameters 1) dynamic
blocking of the parameters \citep{turek2016efficient}, and 2) a custom
blocking scheme were the parameters of each species are blocked
together.

@@

Cross-validation is one of the most fudamental procedures in model
selection, but, because it requires iternativly re-fitting the model,
is computationally intensive \citep{hooten2014guide}. In cross
validation, we exclude a subset of the data ($y_k$) from model
fitting, then use the fitted model to predict $y_k$. The preidtion
error is summarized by coparing the simulated $y_k$ to the true $y_k$.

@@

\section*{Results}
\label{sec:results}

>>>>

First, to improve efficiency, filter over latent states
to calculate model likelihoods in order to limit MCMC sampling to
top-level parameters. We then use two computational approaches to
improve the efficiency of this model 1) dynamic blocking of the
parameters \citep{turek2016efficient}, and 2) a custom MCMC
specification where slice samplers \cite{neal-03} are used for all parameters.

Slice samplers are a class of methods that sample from a target distribution  by using that fact that samples from any distribution can be obtained by sampling uniformly from the area under that distribution's probability density function curve.  The horizontal coordinates of these uniform samples will provide samples from the distribution of interest.  Slice samplers have been shown to perform well in situations where choosing a tuning parameter for a Metropolis algorithm is difficult.

%probably TMI below
When used to sample from the posterior distribution of a univariate parameter, a slice sampler proceeds at each iteration by first choosing a vertical coordinate sampled uniformly between 0 and the height of the density curve at the parameter value from the previous iteration.  Then, a horizontal coordinate is chosen uniformly from the set of all possible parameter values whose density is at least as great as the chosen vertical coordinate.

@@

To improve the efficiency of this model, we first filtered over latent
states to calculate model likelihoods in order to limit MCMC sampling
to top-level parameters. We also vectorized all calculations that
would have require for loops in BUGS or JAGS. We then tried two
approaches to speed sampling of the top-level parameters 1) dynamic
blocking of the parameters \citep{turek2016efficient}, and 2) a custom
blocking scheme where the parameters of each species are blocked
together with adaptive random walk MCMC.

Random walk Metropolis algorithms \citep{metropolis1953equation} sample from the posterior distribution of a parameter of interest by first proposing a new parameter value from a normal proposal distribution centered at the current value, and then deciding whether to accept or reject the proposed value via the Metropolis ratio.  Although common, standard univariate or multivariate normal proposal distributions can prove to be inefficient at sampling for models in which parameters are highly correlated.   Adaptive random walk Metropolis sampling \citep{haario98anadaptive} allows the correlation structure of the posterior distribution of blocks of parameters to be used to produce better proposals.  For an occupancy model where parameters are highly correlated within species, this can provide much more efficient sampling than a non-adaptive Metropolis algorithm.

@@

Cross-validation is one of the most fudamental procedures in model
selection, but, because it requires iternativly re-fitting the model,
is computationally intensive \citep{hooten2014guide}. In cross
validation, we exclude a subset of the data ($y_k$) from model
fitting, then use the fitted model to predict $y_k$. The prediction
error is summarized by comparing the simulated $y_k$ to the true $y_k$.

@@

\section*{Results}
\label{sec:results}
\subsubsection*{Single species, single season occupancy model with
  spatial auto-correlation}

\subsubsection*{Single species, multi season (dynamic) occupancy
  model}
